---
title: "FlightMap"
author: "Alexander Pracht"
date: "10/14/2020"
output: html_document
---
On this page I would like to analyse the current flight schedule of Lufthansa Cargo. I downloaded the data directly from the [Lufthansa Cargo Website](https://lufthansa-cargo.com/network/schedule-routings).

I wanted to check my hypothesis that Lufthansa Cargo offers few flights right now, but will expand their flight schedule from November on.

For visualisation I looked at the following [webpage](https://weiminwang.blog/2015/06/24/use-r-to-plot-flight-routes-on-a-fancy-world-background/).
However some of the functions on this page are already outdated.

```{r, plotting, fig.width=15}
library(ggplot2)
library(maps)
library(rgeos)
library(maptools)
library(ggmap)
library(geosphere)
library(plyr)
library(dplyr)
library(rgdal)
library(data.table)
count.dups <- function(DF){
  DT <- data.table(DF)
  DT[,.N, by = names(DT)]
}

fortify.SpatialLinesDataFrame = function(model, data, ...){
ldply(model@lines, fortify)
}
# importing the data from the csv file
lh_data = read.csv("data/LHcargo_FlightScheduleALL.csv", header=TRUE)

# selecting relevant columns and format dates
lh_data_sliced <-lh_data %>%
  mutate(Mo = ifelse(is.na(Mo), 0, 1),
         Tu = ifelse(is.na(Tu), 0, 1),
         We = ifelse(is.na(We), 0, 1),
         Th = ifelse(is.na(Th), 0, 1),
         Fr = ifelse(is.na(Fr), 0, 1),
         Sa = ifelse(is.na(Sa), 0, 1),
         So = ifelse(is.na(So), 0, 1))%>%
  mutate(sum = Mo+Tu+We+Th+Fr+Sa+So)%>%
  select(DEP,ARR,Start_Op, End_Op, sum)%>%
  mutate(Start_Op = as.Date.character(Start_Op, format = "%d%b%y"), 
         End_Op = as.Date(End_Op, format = "%d%b%y"))


# filter for the current flight schedule as of 15th October 2020
lh_15_10 <- lh_data_sliced%>%
  filter(Start_Op <= "2020-10-15")%>%
    group_by(DEP,ARR)%>%
  summarise(count = sum(sum))

# filter for flight schedule as of November 2020
lh_11 <- lh_data_sliced%>%
  filter(Start_Op <= "2020-11-30", End_Op >= "2020-11-01")%>%
  group_by(DEP,ARR)%>%
  summarise(count = sum(sum))

# formatting and using airport codes to translate to geographical long and lat
airports_dep = read.csv("data/airports.csv", header=FALSE)%>%
  select(V5, V7, V8)

colnames(airports_dep) <- c("DEP", "Lat_Dep", "Long_Dep")

airports_arr = read.csv("data/airports.csv", header=FALSE)%>%
  select(V5, V7, V8)

colnames(airports_arr) <- c("ARR", "Lat_Arr", "Long_Arr")

# include geographical long and lat to the 2 flight schedules
schedule_final_10 <- left_join(lh_15_10, airports_arr)
schedule_final_10 <- left_join(schedule_final_10, airports_dep)
schedule_final_11 <- left_join(lh_11, airports_arr)
schedule_final_11 <- left_join(schedule_final_11, airports_dep)


# delete duplicates to only show 
count_final_10 <- count.dups(schedule_final_10)
count_final_10 <- na.omit(count_final_10)
count_final_11 <- count.dups(schedule_final_11)
count_final_11 <- na.omit(count_final_11)
 
# get worldmap
worldmap = map_data ("world")

# set city lights
library(rgdal) 
urbanareasin <- rgdal::readOGR("data/ne_10m_urban_areas/ne_10m_urban_areas.shp")
urb <- c(geom_polygon(aes(long, lat, group = group),
size = 0.3,
color = "#ffffff",
fill = "#ffffff",
alpha = 1,
data = urbanareasin))

# wrld layer
wrld<-c(geom_polygon(aes(long,lat,group=group), size = 0.1, colour= "#090D2A",
fill="#090D2A", alpha=0.8, data=worldmap))
```

## Lufthansa Cargo Flight Schedule (15th October 2020)

```{r, plotting, fig.width=15}
# October
ggplot() +
wrld +
urb+
geom_segment(data= schedule_final_10, aes(y=Lat_Dep,x=Long_Dep, yend=Lat_Arr, xend=Long_Arr, colour=count), alpha = 0.3)+ 
  scale_color_gradientn(colours = rainbow(20), limits = c(0,166))+
theme(panel.background = element_rect(fill='#00001C',colour='#00001C'), panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
theme(legend.position = c(0,0.4),
legend.justification = c(0,1),
legend.background = element_rect(colour = NA, fill = NA),
legend.key = element_rect(colour = NA, fill = NA, size = 10),
legend.text = element_text(colour='white', size = 20)) +
#guides(fill = guide_legend(keywidth = 20, keyheight = 20)) +
annotate("text",x=max(worldmap$long),y=-60,hjust=.9,size=6,
label=paste("Flights Lufthansa Cargo October"),color="white")
```

## Lufthansa Cargo Flight Schedule (November 2020)

```{r, plotting, fig.width=15}
# November
ggplot() +
wrld +
urb+
geom_segment(data= schedule_final_11, aes(y=Lat_Dep,x=Long_Dep, yend=Lat_Arr, xend=Long_Arr, colour=count), alpha = 0.3)+ 
  scale_color_gradientn(colours = rainbow(20), limits = c(0,166))+
theme(panel.background = element_rect(fill='#00001C',colour='#00001C'), panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
theme(legend.position = c(0,0.4),
legend.justification = c(0,1),
legend.background = element_rect(colour = NA, fill = NA),
legend.key = element_rect(colour = NA, fill = NA, size = 10),
legend.text = element_text(colour='white', size = 20)) +
#guides(fill = guide_legend(keywidth = 20, keyheight = 20)) +
annotate("text",x=max(worldmap$long),y=-60,hjust=.9,size=6,
label=paste("Flights Lufthansa Cargo November"),color="white")

```

What we can see from these graphs is that the flight schedule in October has less flights than the one for November. This means that Lufthansa Cargo believes they can increase their flights and it network, offering more flights to more places.


